---
title: "mergeContacts"
author: "Lucas Carter"
date: "2023-04-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown contains the main functions and surrounding code for merging the contacts extracted from the Lamins Hi-C replicates. There are 6 replicates for each condition and 3 conditions. The Hi-C data was given to us as .hic files, not FASTQs. We were not able to merge the replicates in earlier steps in the analysis as a result. This code is for use on a HPC cluster or cloud. 

## Load Packages

Loads initial packages

```{r, eval = FALSE}

rm(list = ls())

# import packages
library(ggplot2)
library(viridis)
library(dplyr)
library(tidyverse)

```

## mergeContacts() function

mergeContacts() reads in each replicate as a matrix, extracts the bed data from the contact matrices (chrom, pos 1, and pos 2) for use later, removes those positions, and then performs matrix addition to the frequency values in the remaining matrix. After those steps, the positions are added back to the summed matrix, and the final matrix is exported as a tab-delimited table. Function defined below:

```{r, eval = FALSE}
mergeContacts <- function(exp, chrom, res, norm){
  
  # Set directory variables here
  dir.root <- "/projects/b1042/BackmanLab/HiC2/opt/juicer/work/Lamin_HiC/contact_data"
  dir.1 <-paste0(dir.root,"/",exp,"/Rep1/contacts_TopDom/")
  dir.2 <-paste0(dir.root,"/",exp,"/Rep2/contacts_TopDom/")
  dir.3 <-paste0(dir.root,"/",exp,"/Rep3/contacts_TopDom/")
  dir.4 <-paste0(dir.root,"/",exp,"/Rep4/contacts_TopDom/")
  dir.5 <-paste0(dir.root,"/",exp,"/Rep5/contacts_TopDom/")
  dir.6 <-paste0(dir.root,"/",exp,"/Rep6/contacts_TopDom/")
  print(dir.1)
  
  ## Read in table
  m1 <- read.table(file.path(paste0(dir.1, chrom, "-observed_", res,"Kb_",norm,"norm.txt")), sep= "\t", header= F)
  m2 <- read.table(file.path(paste0(dir.2, chrom, "-observed_", res,"Kb_",norm,"norm.txt")), sep= "\t", header= F)
  m3 <- read.table(file.path(paste0(dir.3, chrom, "-observed_", res,"Kb_",norm,"norm.txt")), sep= "\t", header= F)
  m4 <- read.table(file.path(paste0(dir.4, chrom, "-observed_", res,"Kb_",norm,"norm.txt")), sep= "\t", header= F)
  m5 <- read.table(file.path(paste0(dir.5, chrom, "-observed_", res,"Kb_",norm,"norm.txt")), sep= "\t", header= F)
  m6 <- read.table(file.path(paste0(dir.6, chrom, "-observed_", res,"Kb_",norm,"norm.txt")), sep= "\t", header= F)
  
  ## Save positions for later
  ch3 <- m1$V1
  pos1 <- m1$V2
  pos2 <- m1$V3
  
  m1 <- m1[,-c(1:3)]
  m2 <- m2[,-c(1:3)]
  m3 <- m3[,-c(1:3)]
  m4 <- m4[,-c(1:3)]
  m5 <- m5[,-c(1:3)]
  m6 <- m6[,-c(1:3)]
  
  ## Sum matrices
  mat <- m1+m2+m3+m4+m5+m6
  
  ## Add positions back 
  mat <- cbind(ch3, pos1, pos2, mat)
  head(mat, 20)
  
  write.table(mat, file = paste0(dir.root,"/",exp,"/",chrom, "-observed_", res,"Kb_",norm,"norm.txt"), sep = "\t", row.names = F, col.names = F)
  cat(paste('finished processing ',exp,": ",chrom,'\n\n'))
}
```
## Perform mergers on replicates

mergeContacts() is called for each condition in a for-loop below:

```{r, eval = FALSE}

## loop for import
chroms <- c("chr1", "chr2", "chr3", "chr4", "chr5", "chr6", "chr7", "chr8", "chr9", "chr10", "chr11", 
            "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", 
            "chr22", "chrX", "chrY")

for (chrom in 1:length(chroms)){
  
  chrom <- chroms[chrom]
  print(chrom)
  
  mergeContacts(exp="24hrAuxin", chrom, res=25, norm="VC")
}


for (chrom in 1:length(chroms)){
  
  chrom <- chroms[chrom]
  print(chrom)
  
  mergeContacts(exp="untreated", chrom, res=25, norm="VC")
}


for (chrom in 1:length(chroms)){
  
  chrom <- chroms[chrom]
  print(chrom)
  
  mergeContacts(exp="withdraw", chrom, res=25, norm="VC")
}   

```


