---
title: "TAD Analysis Of Merged Contacts Using TopDom"
author: "Lucas Carter"
date: "2023-04-18"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown contains the function for calling the R package, [TopDom](https://github.com/HenrikBengtsson/TopDom). TopDom calls TADs from HiC data. A detailed description of their algorithm can be found in [their paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4838359/). 

## Load Packages

Loads initial packages

```{r, eval = FALSE}

rm(list = ls())

# import packages
library(TopDom)

```

## Define variables

Variables are defined below for each project. We need the condition labels (there are three for this project), the chromosomes that need to be analyzed, the window size, resolution of the Hi-C data, and normalization method (vanilla coverage in this case).

```{r, eval = FALSE}
##------------------------------ values for variables ------------------------------##

## Change these to your condition labels
conds <- c("24hrAuxin", "untreated", "withdraw")
## Change these to the chromosome annotations you're using
chroms = c('1','2','3','4','5','6','7','8','9','10','11','12','13','14','15','16','17','18','19','20','21','22','X')

windowSize = 5 ## TopDom window
res = 25 ## Change to your resolution
norm = "VC" ## Change to your normalization

##-------------------------------values for variables------------------------------##
```

## callTADs() Function

This function runs each condition through a nested for-loop, inputs the contact files, called TopDom, and exports three files containing domain calls and auxillary data for those calls. Function is defined below:

```{r, eval = FALSE}

## Function to loop through and call TADs
callTADs <- function(exp, conds, res, windowSize, norm, chroms){

for (cond in 1:length(conds)){

  cond <- conds[cond] ## Condition labels

  ## Input and output paths on Quest
  input.path = paste0('/projects/b1042/BackmanLab/HiC2/opt/juicer/work/',exp,'/contact_data/',cond,'/merged/')
  output.path = paste0('/projects/b1042/BackmanLab/HiC2/opt/juicer/work/',exp,'/contact_data/',cond,'/merged/domains/')


  for (chrom in chroms) {
    print(chrom)

    input.fname = paste('chr',chrom,'-observed_',res,'Kb_',norm,'norm.txt',sep="")
    output.fname = paste('chr',chrom,'-TADs-window',as.character(windowSize),sep="")

    topDom_file = paste(input.path,input.fname,sep="")
    topDom_outFile = paste(output.path,output.fname,sep="")

    topDom_out <- TopDom(topDom_file, window.size=windowSize, outFile=topDom_outFile)
    }
  }
}

```

Here we call the function:

```{r, eval = FALSE}
callTADs("Lamin_HiC", conds=c("24hrAuxin", "untreated", "withdraw"), res = 25, windowSize = 5, norm = "VC", chroms = c('1','2','3','4','5','6','7','8','9','10','11','12','13','14','15','16','17','18','19','20','21','22','X'))
```

